---
title: "Using stat2rmarkdown Shared Core Utilities"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using stat2rmarkdown Shared Core Utilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(stat2rmarkdown)
```

## Overview

The `stat2rmarkdown` package provides shared utility functions used by the family
of statistical software to R Markdown converter packages:

- `stata2rmarkdown` -- Stata .do/.dta to R Markdown
- `spss2rmarkdown` -- SPSS .sps/.sav to R Markdown
- `jasp2rmarkdown` -- JASP .jasp to R Markdown
- `jamovi2rmarkdown` -- jamovi .omv to R Markdown

The converters do not currently import this package: each carries its own copy
of the equivalent helpers, and `render_jmv_tables()` here is the reference
implementation those copies are kept in step with.

## Utility Functions

### Variable Quoting

When generating R code that references variable names from external data files,
some names may contain spaces, special characters, or start with digits. The
`quote_var()` function handles this:
```{r}
# Safe names are returned as-is
quote_var("age")
quote_var("score_total")

# Names that need quoting get backticks
quote_var("income (USD)")
quote_var("2nd_measurement")
```

### Null Coalescing

The `%||%` operator provides a concise way to supply default values:
```{r}
user_title <- NULL
title <- user_title %||% "Untitled Analysis"
title

# Non-NULL values pass through
user_title <- "My Study"
title <- user_title %||% "Untitled Analysis"
title
```

### Safe Numeric Conversion

Data from statistical packages often arrives as factors or labelled vectors.
`safe_as_numeric()` handles the conversion safely:
```{r}
# Factor to numeric (via character, preserving labels)
safe_as_numeric(factor(c("1.5", "2.0", "3.5")))

# Character to numeric (NAs for non-numeric)
safe_as_numeric(c("10", "abc", "30"))
```

### Variable Validation

Before running analyses, check that expected variables exist in the data:
```{r, error=TRUE}
df <- data.frame(Age = 1:5, Score = rnorm(5), Group = rep("A", 5))

# Case-insensitive by default (handles Stata/SPSS conventions)
validate_vars(df, c("age", "SCORE"))

# Fails with a clear error for missing variables
validate_vars(df, c("age", "nonexistent"))
```

### Logging

Track conversion progress with timestamped messages:
```{r}
log_message("Starting conversion")
log_message("Variable 'x' not found", level = "WARN")
```

### Output Directory Management

Create organized, timestamped output directories:
```{r, eval=FALSE}
dir <- create_output_dir("stata_analysis")
# Creates: output/2026-03-28-14-30-00-stata_analysis/
```

## R Markdown Skeleton

The skeleton functions generate the building blocks of an R Markdown document.

### YAML Header

```{r}
yaml <- generate_yaml_header(
 title = "Stata to R Conversion: my_study",
 author = "Generated by stata2rmarkdown"
)
cat(yaml)
```

### Setup Chunk

```{r}
setup_chunk <- generate_setup_chunk(
  packages = c("haven", "dplyr", "fixest", "modelsummary"),
  extra_code = "setFixest_notes(FALSE)\nmodels <- list()"
)
cat(setup_chunk)
```

### Session Footer

```{r}
footer <- generate_session_footer(
  tool_name = "stata2rmarkdown",
  repo_url = "https://github.com/giladfeldman/STATA2Rmarkdown"
)
cat(footer)
```

### Assembling a Complete Document

A converter package typically combines these pieces with its own analysis
sections:

```{r, eval=FALSE}
rmd <- paste(
  generate_yaml_header("My Analysis", author = "Generated by spss2rmarkdown"),
  generate_setup_chunk(packages = c("haven", "dplyr", "psych")),
  "\n# Analysis\n\n... analysis chunks here ...\n",
  generate_session_footer("spss2rmarkdown"),
  sep = "\n"
)
writeLines(rmd, "output/my_analysis.Rmd")
```
